\chapter{Partial Reconfiguration Scripts}

The Partial Reconfiguration workflow, currently (early 2018) is not fully integrated at the Vivado GUI.
In newer versions a designer may use the GUI to create an HDL-only design, but the use if IP Integrator
is still not supported. For now, only the initial design can be created in Vivado GUI.
After floorplanning, the synthesized result will be saved in a \gls{dcp} and the main P.R. workflow
has to be done with the help of TCL scripts. 

Xilinx has implemented a sample design that uses
a single \gls{rm} for implementing an imaging filter. It is an integrated embedded system,
that accepts a video stream from a HDMI capable daughtercard connected via the FMC,
it performs the filtering and it outputs it to an HDMI capable output. It is implemented and
documented in both ISE (see \cite{xapp1159}) and Vivado (see \cite{xapp1231}).

From this application, the core TCL scripts were extracted, and the client script
was rewritten to match our application.

All scripts were created in GNU/Linux and use the UNIX directory structure.
They will not work in a Microsoft Windows environment.

\section{Creating static design Device Checkpoint (DCP)}
This script, run from within Vivado with the synthesized design open, 
will replace the instantiated accelerators with black-boxes and save
the result to a \gls{dcp}.

\lstinputlisting[style=basic,language=tcl]{scripts/gen_dcp.tcl}

\section{Generating the Project Files}
After static design floorplanning and synthesis, a project file must be
created for each \gls{rm} version of each \gls{rp}.
These project files are a requirement for Xilinx TCL scripts.

\lstinputlisting[style=basic,language=bash]{scripts/prepare.sh}

\section{TCL Client Script}
This script is the client to Xilinx partial reconfiguration TCL scripts.
It has to be modified to match the FPGA and hardware design of the designer.
It was made to be as flexible as possible for future adaptation to other projects.

\lstinputlisting[style=basic,language=tcl]{scripts/design.tcl}

\section{Partial Bitstream Manipulation}
The Xilinx TCL scripts will generate a large number of files, 
including all partial bitstreams. This script will install them to our software framework,
after first renaming according to the naming conventions set by the system library
and compressing them with the XZ compresison tool.

\lstinputlisting[style=basic,language=bash]{scripts/install.sh}

